---
title: "PA2"
author: "R. Omlin"
date: "25. June 2018"
output: html_document
---

# Storm events and its harmful impact on human and economic

## Synopsis

## Data Processing
### Preparing the Environment
This script has been created on a macOS High Sierra machine (64-bit). RStudio runs on version 1.1.453 and R on version 3.3.2 (64-bit).
First i'm gonna make sure, that your machine uses the same language setting as there could be problems when running the script (especially when dealing with Weekdays). If you're on a windows machine, replace "en_US.UTF-8" with "English", Linux should work ok.
```{r localize}
Sys.setlocale("LC_ALL","en_US.UTF-8")
Sys.setenv(TZ="Europe/Zurich")
```

I had also the need to set the timezone to avoid error messages (i think this can be adapted to your needs).
```{r timezone}
Sys.setenv(TZ="Europe/Zurich")
```
I asume you already installed 'ggplot2' otherwise install it with 'install.packages("ggplot2")' and load it with.
```{r ggplots}
library(ggplot2)
```

### Loading the data
Now the environment should be ready for the rest of the script and we can start downloading the data from the internet using this Chunk.
```{r downloading}
datafilename <- "StormData.csv.bz2" 
datasource <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2FStormData.csv.bz2"
if(!file.exists(datafilename))
{
  download.file(datasource, datafilename, mode="wb")
}
```

This will download the activity.zip file into the working directory. The url used is stored in the variable datasource.
While the file is zipped, it will be unzipped in a variable called temp and then read with the 'read.csv()' function in a data frame named 'df'
```{r loading, cache=TRUE}
df <- read.csv(datafilename)
```

### Processing the data

## Results
